---
title: "Homework 1"
subtitle: "Stat 253"
author: "Mohammed Sohail Khan "
date: 14th Sep 2025
format:
  html:
    toc: true
    toc-depth: 2
    embed-resources: true
    fig-width: 5
    fig-height: 3
execute:
  error: true
  warning: false
knitr:
  opts_chunk:
    collapse: true
    message: false
---


# Notes 

(@) Please **update the author name** to your name! 
(@) This document is for your answers only. Do NOT write out the homework questions in this document. Writing out the questions makes it very challenging to grade the homework. (You can create a separate document for your own records if you wish.) 
(@) Besides adding your answers/code, please DO NOT make any deletions or changes to the structure of this template! 



# Setup 

Put any code you need to load packages and load/prep data here. 

```{r}
library(tidyverse)
library(tidymodels)
library(rattle)

data("weatherAUS")
glimpse(weatherAUS)

view(weatherAUS)
```


# Exercises



## Data Context



## Exercise 1



### Part a
filtering 6 rows namely : Hobart, Adelaide, Canberra, Brisbane, Melbourne, and Sydney. 
Mutating(in this case writing over) temp9am values to be farenheit
and then selecting the relevant 6 columns
```{r}
# I used weather_data dataset as I have created a weather dataset in my other class 
weather_data <-weatherAUS %>% 
  filter(Location %in% c("Hobart", "Adelaide", "Canberra", "Brisbane", "Melbourne", "Sydney") ) %>% 
  mutate(Temp9am = Temp9am *1.8 + 32, Temp3pm = Temp3pm * 1.8 + 32) %>% 
  select(Temp3pm, Location, WindSpeed9am, Humidity9am, Pressure9am, 
      Temp9am)
  

```

```{r}
# Checking it out
dim(weather_data)
head(weather_data)
```


### Part b
Arranging it in desending order and then looking at top 3 using head function proving 3 as an argument

```{r}
top_3_hottest <- weather_data %>% 
  arrange(desc(Temp3pm)) %>% 
  head(3)

```

```{r}
# checking it out 
top_3_hottest

```

### Part c
```{r}
temp3pm_6locs <- weather_data %>% 
  group_by(Location) %>% # makes mini groups of every location and then takes mean of the group  based on temp3pm
  summarise(mean(Temp3pm, na.rm = TRUE))
  
```

```{r}
# lets check it out 
temp3pm_6locs

```


### Part d
```{r}
ggplot(weather_data, aes(x = Temp3pm, fill = Location)) + 
  geom_density(alpha = 0.5) # alpha adds translucency to see overlap
  
```
```{r}
# facet_wrap() -- basically splits one plot into smaller sub plots based on similar values in a column 
ggplot(weather_data, aes(x = Temp9am, y = Temp3pm,color = Humidity9am)) + geom_point() + facet_wrap(~Location) +
scale_color_gradient(low ="#56B1F7" , high = "#132B43")

```



### Part e



## Exercise 2



### Part a



### Part b



### Part c



### Part d



### Part e



### Part f



## Exercise 3



### Part a



### Part b



### Part c



## Exercise 4



### Part a



### Part b



### Part c



### Part d



### Part e



## Exercise 5



## Exercise 6




# Done!

- Render your QMD to HTML.
- Check that HTML to make sure there are no errors or formatting issues. 
- Submit the HTML on Moodle.


